---
title: "Data Harmony"
author: "Brian Yandell"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r}
library(tidyverse)
library(readxl)
library(readr)
library(foundr) # github/byandell/foundr
```

```{r}
for(filename in list.files("R", full.names = TRUE)) {
  source(filename)
}
```

# Harmonize Data

Data for this repository are identified by `data/source.csv`,
which is saved with the repo.

```{r}
rootdir <- "/mnt/researchdrive/adattie/General/founder_diet_study"
rawdir <- file.path(rootdir, "RawData")
harmonizeddir <- file.path(rootdir, "HarmonizedData")
```

```{r}
setwd(harmonizeddir)
```

```{r}
links <- read.csv(file.path("data", "source.csv"), fill = TRUE) %>%
  mutate(address = file.path(rawdir, address))
```

## Annotation

The experiment annotation file relates the `number` (identifier for `animal`) to the `sex` and `diet`. Needed for `LivRna`.

```{r}
annotfile <- linkpath("annot", links)
excel_sheets(annotfile)
annot <- read_excel(annotfile) %>%
  mutate(diet = ifelse(as.character(diet_no) == "200339", "HC_LF", "HF_LC"))
```


## Physio: Physiological traits

```{r}
harmonize("Physio", links, PhysioHarmony, condition_name = "diet")
```

## PlaMet: Plasma metabolites

```{r warning = FALSE}
harmonize("PlaMet0", links, MetHarmony, annot, skiprows = 4, condition_name = "diet")
```

## PlaMet120: Plasma metabolites at 120 minutes

```{r warning = FALSE}
harmonize("PlaMet120", links, MetHarmony, annot, skiprows = 3, condition_name = "diet")
```

## LivMet: Liver metabolites

```{r warning = FALSE}
harmonize("LivMet", links, MetHarmony, annot, skiprows = 3, condition_name = "diet")
```

## LivRna: Liver mRNA expression

There is a separate liver annotation file, which relates `ENTREZID` to `SYMBOL`.

```{r}
liverAnnot <- linkpath("liver_annot", links)
liverAnnot <- read_csv(liverAnnot)
```

```{r warning = FALSE}
harmonize("LivRna", links, LivRnaHarmony, annot, liverAnnot, condition_name = "diet")
```

## Enrich: Plasma enrichment over time

```{r}
harmonize("Enrich", links, EnrichHarmony, annot, condition_name = "diet")
```

## Module: WGCNA module eigentraits

See <WGCNA.Rmd> for construction of `traitModule`.

```{r}
traitModule <- readRDS("traitModule.rds")
```

```{r}
harmonize("Module", links, foundr::moduleHarmony,
          traitModule, condition_name = "diet")
```

```{r}
bind_traits(c("Physio","PlaMet0","PlaMet120","LivMet","LivRna","Module","Enrich"))
```
