---
title: "Founder Diet Study Data"
author: "Brian Yandell"
date: "2022-12-05"
output: html_document
---

```{r setup, include=FALSE, echo = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                      comment = "", fig.width = 7, fig.height = 7)
```
 
```{r}
library(tidyverse)
library(readxl)
library(readr)
```

Data for this repository are identified by `data/source.csv`,
which is not saved with the repo.

```{r}
links <- read_csv(file.path("data", "source.csv"))
linkpath <- function(dataset, datadir = "data") {
  filename <- (links %>% filter(object == dataset))$address
  if(datadir != "")
  filename <- file.path(datadir, filename)
  filename
}
```

Data comes from Google Drive <`r linkpath("google")`>.
See the QC Analysis Report on liver RNAseq data (in zip file under folder `Liver RNAseq`).

```{r}
broomit <- function(traitsdata) {
  map(
    split(traitsdata, traitsdata$trait),
    function(traitdata) {
      ct <- distinct(traitdata, strain, sex, diet)
      ct <- count(ct, sex, diet)
      if(nrow(ct) < 4 | min(ct$n) < 2) {
        # Do nothing if any combination of sex, diet is empty.
        return(NULL)
      }
      form <- formula(value ~ strain * sex * diet)
      fit <- stats::lm(form, traitdata)
      as.tibble(data.frame(
        rawSD = sd(traitdata$value, na.rm = TRUE),
        broom::tidy(stats::drop1(fit, fit, test = "F")) %>%
          filter(grepl("strain:", term)) %>%
          select(term, p.value) %>%
          pivot_wider(names_from = "term",
                      values_from = "p.value")))
      }) %>%
  bind_rows(.id = "trait")
}
```

# Mouse annotations

```{r}
annotfile <- linkpath("annot")
excel_sheets(annotfile)
annot <- read_excel(annotfile) %>%
  mutate(diet = ifelse(as.character(diet_no) == "200339", "HC_LF", "HF_LC"))
```

```{r}
annot %>%
  distinct(diet, diet_no, diet_description)
```

```{r}
annot %>%
  count(strain, sex, diet) %>%
  unite("sex_diet", sex, diet) %>%
  pivot_wider(names_from = "sex_diet", values_from = "n")
```

# Physiological trait counts by strain, sex, diet

```{r}
physiofile <- linkpath("physio")
excel_sheets(physiofile)
physiodef <- read_excel(physiofile, sheet = 1, skip = 1) %>%
  unite("comment", -(1:3)) %>%
  mutate(comment = str_remove(str_replace_all(
    str_remove_all(comment, "NA"), "_+", "_"), "^_"))
physio <- read_excel(physiofile, sheet = 2) %>%
  pivot_longer(-(1:4), names_to = "trait", values_to = "value") %>%
  mutate(diet = ifelse(as.character(diet) == "200339", "HC_LF", "HF_LC"))
```

There are `r nrow(distinct(physio, trait))` distinct traits.

Count of mice per strain, sex and diet. All combinations have at least some measurements.

```{r}
physio %>%
  filter(!is.na(value)) %>%
  distinct(strain, number, sex, diet) %>%
  count(strain, sex, diet) %>%
  unite("sex_diet", sex, diet) %>%
  pivot_wider(names_from = "sex_diet", values_from = "n")
```

Percent missing values across traits

```{r}
physio %>%
  filter(!is.na(value)) %>%
  count(strain, sex, diet) %>%
  unite("sex_diet", sex, diet) %>%
  mutate(n = round(100 * (1 - n / max(n)), 2)) %>%
  pivot_wider(names_from = "sex_diet", values_from = "n")
```

```{r}
saveRDS(physio, "physio.rds")
```

## Physiological summaries

```{r}
physiosum <- broomit(physio)
rm(physio)
```

# Plasma metabolite abundance data

```{r}
plasmafile <- linkpath("plasma")
excel_sheets(plasmafile)
removed <- read_excel(plasmafile, sheet = 1, skip = 4) %>%
  pivot_longer(-(1:4), names_to = "mousecode", values_to = "value") %>%
  mutate(strain = str_remove(str_extract(mousecode, "^.*-"), "-"),
         number = str_extract(str_remove(mousecode, "^.*-"), "^[0-9]+"),
         sex = str_extract(str_remove(mousecode, "^.*-[0-9]+_"), "^[MF]"),
         diet = str_remove(str_remove(mousecode, "^.*-[0-9]+_"), "^[MF]_")) %>%
  rename(trait = "compound")
plasma <- read_excel(plasmafile, sheet = 1, skip = 4) %>%
  pivot_longer(-(1:4), names_to = "mousecode", values_to = "value") %>%
  mutate(strain = str_remove(str_extract(mousecode, "^.*-"), "-"),
         number = str_extract(str_remove(mousecode, "^.*-"), "^[0-9]+"),
         sex = str_extract(str_remove(mousecode, "^.*-[0-9]+_"), "^[MF]"),
         diet = str_remove(str_remove(mousecode, "^.*-[0-9]+_"), "^[MF]_")) %>%
  rename(trait = "compound")
```

Removed plasma data appears to be identical.

```{r}
removed_comp <- (removed %>%
  filter(!is.na(value)) %>%
  count(trait))$trait
plasma_comp <- (plasma %>%
  filter(!is.na(value)) %>%
  count(trait))$trait
m <- match(removed_comp, plasma_comp)
sum(is.na(m))
```

```{r}
plasmar <- full_join(
  plasma %>% select(strain, number, sex, diet, trait, value),
  removed %>% select(strain, number, sex, diet, trait, value),
  by = c("strain", "number", "sex", "diet", "trait"),
  suffix = c(".p", ".r"))
summary(plasmar$value.p - plasmar$value.r)
```

There are missing data for B6 mice across 3 of the 4 sex:diet combinations.

```{r}
plasma %>%
  filter(!is.na(value)) %>%
  distinct(strain, number, sex, diet) %>%
  count(strain, sex, diet) %>%
  unite("sex_diet", sex, diet) %>%
  pivot_wider(names_from = "sex_diet", values_from = "n")
```

```{r}
saveRDS(plasma, "plasma.rds")
```

## Plasma summaries

```{r}
plasmasum <- broomit(plasma)
plasmaorder <- distinct(plasma, trait)$trait
rm(plasma)
m <- match(plasmaorder, plasmasum$trait)
plasmasum <- plasmasum[m,]
```

# Liver RNAseq

Note that annotation file relates the number to the sex and diet.

```{r}
liverfile <- linkpath("liver")
liver <- read_csv(liverfile) %>%
  rename(trait = "...1") %>%
  mutate(trait = as.character(trait)) %>%
  pivot_longer(-1, names_to = "strain_number", values_to = "value") %>%
  separate(strain_number, sep = "_", into = c("strain","number")) %>%
  mutate(strain = ifelse(strain == "A129", "129", strain))
```

```{r}
liver_annot <- linkpath("liver_annot")
annot2 <- read_csv(liver_annot)
```

```{r}
liver <- left_join(
  liver,
  annot %>% mutate(number = as.character(number)),
  by = c("strain", "number"))
```

Change trait name to `SYMBOL_ENTREZID`. Note that some mRNA have multiple entries, so we cannot just use the `SYMBOL`.

```{r}
liver <- left_join(
  liver %>% rename(ENTREZID = "trait"),
  annot2[,-1] %>%
    mutate(ENTREZID = as.character(ENTREZID)),
  by = "ENTREZID") %>%
  rename(trait = "SYMBOL") %>%
  mutate(trait = paste(trait, ENTREZID, sep = "_")) %>%
  select(-ENTREZID) %>%
  select(trait, everything())
```

## Reduce to ones with few small values

```{r}
liver <- liver %>%
  filter(value > min(liver$value))
```

Percent missing values

```{r}
liver %>%
  filter(!is.na(value)) %>%
  count(strain, sex, diet) %>%
  unite("sex_diet", sex, diet) %>%
  mutate(n = round(100 * (1 - n / max(n)), 2)) %>%
  pivot_wider(names_from = "sex_diet", values_from = "n")
```

```{r}
saveRDS(liver, "liver.rds")
```

## Liver summaries

```{r}
liversum <- broomit(liver)
rm(liver)
liverorder <- unite(annot2, trait, 3:2)$trait
liverorder <- liverorder[liverorder %in% liversum$trait]
m <- match(liverorder, liversum$trait)
liversum <- liversum[m,]
```

# Summaries

See computations above for `physiosum`, `plasmasum` and `liversum` using `broomit`.

```{r}
traitsum <- bind_rows(
  physio = physiosum,
  plasma = plasmasum,
  liver = liversum,
  .id = "datatype")
saveRDS(traitsum, "traitsum.rds")
```

```{r}
write_csv(traitsum, "trait_summary.csv")
```
