---
title: "Founder Diet Study Data"
author: "Brian Yandell"
date: "2022-12-05"
output: html_document
---

```{r setup, include=FALSE, echo = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                      comment = "", fig.width = 7, fig.height = 7)
```
 
See file `X202SC22070248-Z01-F001_Report.html` for QC Analysis Report on liver RNAseq data.

```{r}
library(tidyverse)
library(readxl)
library(readr)
```

# Mouse annotations

```{r}
excel_sheets("data/mouse annotations for founder diet study.xlsx")
annot <- read_excel("data/mouse annotations for founder diet study.xlsx") %>%
  mutate(diet = ifelse(as.character(diet_no) == "200339", "HC_LF", "HF_LC"))
```

```{r}
annot %>%
  distinct(diet, diet_no, diet_description)
```

```{r}
annot %>%
  count(strain, sex, diet) %>%
  unite("sex_diet", sex, diet) %>%
  pivot_wider(names_from = "sex_diet", values_from = "n")
```

# Physiological trait counts by strain, sex, diet

```{r}
excel_sheets("data/Final raw and computed in vivo founder data for Mark for correlations.xlsx")
physiodef <- read_excel("data/Final raw and computed in vivo founder data for Mark for correlations.xlsx",
                     sheet = 1, skip = 1) %>%
  unite("comment", -(1:3)) %>%
  mutate(comment = str_remove(str_replace_all(
    str_remove_all(comment, "NA"), "_+", "_"), "^_"))
physio <- read_excel("data/Final raw and computed in vivo founder data for Mark for correlations.xlsx",
                     sheet = 2) %>%
  pivot_longer(-(1:4), names_to = "trait", values_to = "value") %>%
  mutate(diet = ifelse(as.character(diet) == "200339", "HC_LF", "HF_LC"))
```

There are `r nrow(distinct(physio, trait))` distinct traits.

Count of mice per strain, sex and diet. All combinations have at least some measurements.

```{r}
physio %>%
  filter(!is.na(value)) %>%
  distinct(strain, number, sex, diet) %>%
  count(strain, sex, diet) %>%
  unite("sex_diet", sex, diet) %>%
  pivot_wider(names_from = "sex_diet", values_from = "n")
```

Percent missing values across traits

```{r}
physio %>%
  filter(!is.na(value)) %>%
  count(strain, sex, diet) %>%
  unite("sex_diet", sex, diet) %>%
  mutate(n = round(100 * (1 - n / max(n)), 2)) %>%
  pivot_wider(names_from = "sex_diet", values_from = "n")
```

```{r}
saveRDS(physio, "physio.rds")
```

# Plasma metabolite abundance data

```{r}
excel_sheets("data/15Aug2022_Founder_T0_metabolomics_combined_unique_isobars_removed.xlsx")
removed <- read_excel("data/15Aug2022_Founder_T0_metabolomics_combined_unique_isobars_removed.xlsx",
                     sheet = 1, skip = 4) %>%
  pivot_longer(-(1:4), names_to = "mousecode", values_to = "value") %>%
  mutate(strain = str_remove(str_extract(mousecode, "^.*-"), "-"),
         number = str_extract(str_remove(mousecode, "^.*-"), "^[0-9]+"),
         sex = str_extract(str_remove(mousecode, "^.*-[0-9]+_"), "^[MF]"),
         diet = str_remove(str_remove(mousecode, "^.*-[0-9]+_"), "^[MF]_")) %>%
  rename(trait = "compound")
plasma <- read_excel("data/15Aug2022_Founder_T0_metabolomics_combined_unique_isobars_removed.xlsx",
                     sheet = 1, skip = 4) %>%
  pivot_longer(-(1:4), names_to = "mousecode", values_to = "value") %>%
  mutate(strain = str_remove(str_extract(mousecode, "^.*-"), "-"),
         number = str_extract(str_remove(mousecode, "^.*-"), "^[0-9]+"),
         sex = str_extract(str_remove(mousecode, "^.*-[0-9]+_"), "^[MF]"),
         diet = str_remove(str_remove(mousecode, "^.*-[0-9]+_"), "^[MF]_")) %>%
  rename(trait = "compound")
```

Removed plasma data appears to be identical.

```{r}
removed_comp <- (removed %>%
  filter(!is.na(value)) %>%
  count(trait))$trait
plasma_comp <- (plasma %>%
  filter(!is.na(value)) %>%
  count(trait))$trait
m <- match(removed_comp, plasma_comp)
sum(is.na(m))
```

```{r}
plasmar <- full_join(
  plasma %>% select(strain, number, sex, diet, trait, value),
  removed %>% select(strain, number, sex, diet, trait, value),
  by = c("strain", "number", "sex", "diet", "trait"),
  suffix = c(".p", ".r"))
summary(plasmar$value.p - plasmar$value.r)
```

There are missing data for B6 mice across 3 of the 4 sex:diet combinations.

```{r}
plasma %>%
  filter(!is.na(value)) %>%
  distinct(strain, number, sex, diet) %>%
  count(strain, sex, diet) %>%
  unite("sex_diet", sex, diet) %>%
  pivot_wider(names_from = "sex_diet", values_from = "n")
```

```{r}
saveRDS(plasma, "plasma.rds")
```

# Liver RNAseq

Note that annotation file relates the number to the sex and diet.

```{r}
liver <- read_csv("data/normalizedcount.csv") %>%
  rename(trait = "...1") %>%
  mutate(trait = as.character(trait)) %>%
  pivot_longer(-1, names_to = "strain_number", values_to = "value") %>%
  separate(strain_number, sep = "_", into = c("strain","number"))
```

```{r}
liver <- left_join(
  liver,
  annot %>% mutate(number = as.character(number),
                   strain = ifelse(strain == "129", "A129", strain)),
  by = c("strain", "number"))
```

```{r}
saveRDS(liver, "liver.rds")
```

# Analysis

```{r}
broomit <- function(traitsdata) {
  map(
    split(traitsdata, traitsdata$trait),
    function(traitdata) {
      form <- formula(value ~ strain * sex * diet)
      fit <- stats::lm(form, traitdata)
      as.tibble(data.frame(
        rawSD = sd(traitdata$value, na.rm = TRUE),
        broom::tidy(stats::drop1(fit, fit, test = "F")) %>%
          filter(grepl("strain:", term)) %>%
          select(term, p.value) %>%
          pivot_wider(names_from = "term",
                      values_from = "p.value")))
      }) %>%
  bind_rows(.id = "trait")
}
```

```{r}
physio <- readRDS("physio.rds")
```

```{r}
physiosum <- broomit(physio)
```

```{r}
head(physiosum)
```

```{r}
plasmasum <- broomit(readRDS("plasma.rds"))
```

```{r}
liversum <- broomit(readRDS("liver.rds"))
```

```{r}
saveRDS(
  bind_rows(
    physio = physiosum,
    plasma = plasmasum,
    liver = liversum,
    .id = "datatype"),
  "traitsum.rds")
```

