---
title: "Founder Diet Study Data"
author: "Brian Yandell"
date: "2022-12-05"
output: html_document
params:
  dataset:
    label: "Measurement Set"
    value: plasma
    input: select
    choices: [physio, plasma, liver]
  traits:
    label: "Number of traits"
    value: 3
    input: slider
    min: 2
    max: 6
    
---

```{r setup, include=FALSE, echo = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                      comment = "", fig.width = 7, fig.height = 7)
```

```{r}
library(tidyverse)
library(readxl)
library(readr)
```

```{r}
meas <- readRDS("traits.rds") %>%
  filter(datatype %in% params$dataset)
measum <- readRDS("traitsum.rds") %>%
  filter(datatype %in% params$dataset) %>%
  arrange(strain.sex.diet)
```

```{r}
traits <- measum$trait[seq_len(as.integer(params$traits))]
meas <- meas %>% 
  filter(trait %in% traits)
```

Here we pick out 6 compounds with the lowest absolute value and compares sex, diet and strain.

```{r warnings = FALSE}
ggplot(meas %>%
         unite(sex_diet, sex, diet)) +
  aes(sex_diet, value, col = sex_diet) +
  geom_jitter() +
  facet_grid(trait ~ strain, scales = "free_y") +
  theme(legend.position = "none", axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
```

```{r}
dat <- meas %>%
  filter(trait %in% traits[1:2]) %>%
  select(strain, number, sex, diet, trait, value) %>%
  pivot_wider(names_from = "trait", values_from = "value")
load("data/CCcolors.RData")
plotly::ggplotly(
  ggplot(dat) +
    aes(.data[[traits[1]]], .data[[traits[2]]], color = strain) +
    geom_smooth(method = "lm", se = FALSE) +
    geom_point(size = 2) +
    scale_color_manual(values = CCcolors) +
    facet_grid(sex ~ diet, scales = "free")) 
```

## More than two pairs

This works for 3 but gets crowded for more

```{r}
traitpairs <- as.data.frame(combn(traits, 2))
dat <- 
  bind_rows(
    map(
      traitpairs,
      function(x) {
        meas %>%
          filter(trait %in% x) %>%
          mutate(trait = c("left","right")[match(trait, x)]) %>%
          select(strain, number, sex, diet, trait, value) %>%
          pivot_wider(names_from = "trait", values_from = "value")
      }),
    .id = "pair") %>%
  unite(sex_diet, sex, diet) %>%
  mutate(pair1 = as.matrix(traitpairs)[1, pair],
         pair2 = as.matrix(traitpairs)[2, pair]) %>%
  select(pair1, pair2, everything())

plotly::ggplotly(
  ggplot(dat) +
    aes(left, right, color = strain) +
    geom_smooth(method = "lm", se = FALSE) +
    geom_point(size = 2) +
    scale_color_manual(values = CCcolors) +
    facet_grid(pair2 + pair1 ~ sex_diet, scales = "free"))
```

```{r}
traitpairs2 <- traitpairs %>%
  mutate(across(
    everything(), 
    function(x) {
      c(paste(x, collapse = ":"),
      paste(rev(x), collapse = ":"))
      })) %>%
  unlist()
```

```{r}
dat <- 
  map(
    traitpairs2,
    function(x) {
      x <- str_split(x, ":")[[1]]
      out <- meas %>%
        filter(trait %in% x) %>%
        mutate(trait = c(x[1],x[2])[match(trait, x)]) %>%
        select(strain, number, sex, diet, trait, value) %>%
        pivot_wider(names_from = "trait", values_from = "value") %>%
        unite(sex_diet, sex, diet)
      ggplot(out) +
        aes(.data[[x[1]]], .data[[x[2]]], color = strain) +
        geom_smooth(method = "lm", se = FALSE, formula = "y ~ x") +
        geom_point(size = 2) +
        scale_color_manual(values = CCcolors) +
        facet_grid(. ~ sex_diet) +
        theme(legend.position = "none")
    })
```


```{r}
tmp <- names(dat)[1:3]
patchwork::wrap_plots(dat[tmp], nrow = length(tmp))
```

## Other ideas

Replaced variability in order (which does not really work) with
p_overall something like `broom::tidy(stats::anova(fitfull, fitred))`
where fitfull has model 

`trait ~ strain * sex * diet`

and fitred has model

`trait ~ strain * se + sex * diet`

That is, we remove overall strain effect but combine `strain:sex`, `strain:diet` and `strain:sex:diet`.

```{r}
tmp <- meas %>% filter(trait == traits[1])
formful <- formula(value ~ strain * sex * diet)
fitful <- stats::lm(formful, tmp)
formred <- formula(value ~ strain * sex + sex * diet)
fitred <- stats::lm(formred, tmp)
      
(broom::tidy(stats::anova(fitred, fitful)) %>%
  select(p.value))[2,]$p.value
```

```{r}
rr <- resid(fitred)
```

# Residual from Reduced fit for Dimension Reduction

```{r}
physio <- readRDS("physio.rds")
```

```{r}
out <- residit(physio)
```

Spearman rank correlation of residuals after removing `strain*sex` and `diet*sex` effects.
That is, looking at correlation due to `strain*diet` and `strain*diet*sex` effects only.

```{r}
cors <- out %>%
  select(-value) %>%
  pivot_wider(names_from = "trait", values_from = "residred") %>%
  select(-strain, -number, -sex, -diet) %>%
  cor(use = "pairwise", method = "spearman")
```

May want to eliminate some with low correlation

```{r}
summary(apply(cors, 1, function(x) max(x[x<1], na.rm = TRUE)))
```

Or otherwise use ideas like WCGNA / clustering to find similar traits.
