---
title: "Founder Diet Study Correlations"
author: "Brian Yandell"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
params:
  dataset:
    label: "Measurement Set"
    value: plasma
    input: select
    choices: [physio, plasma, liver]
  traits:
    label: "Number of traits"
    value: 3
    input: slider
    min: 2
    max: 6
    
---

```{r setup, include=FALSE, echo = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                      comment = "", fig.height = 7, fig.width = 7)
```

```{r}
library(tidyverse)
library(readxl)
library(readr)
library(foundr) # github/byandell/foundr
```

# Compare raw and adjusted correlations

```{r}
plasma <- readRDS("traits.rds") %>%
  filter(datatype == "plasma")
```

```{r}
if(file.exists(filerds <- "plasmasignal.rds")) {
  out <- readRDS(filerds)
} else {
  out <- partition(plasma)
  saveRDS(out, filerds)
}
```

```{r}
outw <- out %>%
  select(-mean) %>%
  pivot_wider(names_from = "trait", values_from = "signal")
write.csv(outw, "plasmasignal.csv", row.names = FALSE)
```

```{r}
outv <- 
  left_join(
    plasma %>% select(-datatype),
    out,
    by = c("trait", "strain", "sex"))
```

```{r}
corval <- 
  bind_rows(
    map(
    split(outv, outv$trait),
    function(x) {
      terms <- c("signal","mean")
      as.data.frame(cor(x[,"value"], x[,terms], use = "pair"))
    }),
    .id = "trait")
apply(corval[,-1], 2, summary)
```

Pattern below makes sense because `mean` captures most of `value`, so their correlations range up toward 1. The `signal` has no clear pattern of correlation with `value`. However, `signal` is that part of the `value` consists of the terms `strain:diet` and `strain:diet:sex`, the important part of the signal.

```{r}
GGally::ggpairs(corval[,-1])
```


Spearman rank correlation of residuals after removing `strain*sex` and `diet*sex` effects.
That is, looking at correlation due to `strain*diet` and `strain*diet*sex` effects only.

```{r}
if(file.exists(filerds <- "corsraw.rds")) {
  corsraw <- readRDS(filerds)
} else {
  corsraw <- abscorcalc(plasma)
  saveRDS(corsraw, filerds)  
}
if(file.exists(filerds <- "cors.rds")) {
  cors <- readRDS(filerds)
} else {
  cors <- abscorcalc(out %>% select(-mean) %>% rename(value = "signal"))
  saveRDS(cors, filerds)  
}
```

```{r}
traitsum <- readRDS("traitsum.rds")
tmp <- (traitsum %>% filter(datatype == "plasma"))$p_signal
pvalr <- matrix(tmp, byrow = TRUE, nrow = length(tmp), ncol = length(tmp))
pvalr <- pvalr[upper.tri(pvalr)]
pvalc <- matrix(tmp, byrow = FALSE, nrow = length(tmp), ncol = length(tmp))
pvalc <- pvalc[upper.tri(pvalc)]
```

```{r}
dat <- data.frame(raw = corsraw[upper.tri(corsraw)],
                  adj = cors[upper.tri(cors)]) %>%
  mutate(adjgrp = 1 + pmin(3, floor(raw * 4)),
         pvalr = -log10(pvalr),
         pvalc = -log10(pvalc),
         logpval = pmin(5, pmax(pvalr, pvalc)),
         logpmax = pmin(5, pmax(pvalr, pvalc)))
```

Each point is a pair of traits showing their raw and adjusted (signal only) correlations. Coloring is by the most significant `p_signal`
for the two traits being correlated. This shows essentially no correlation between  until the raw are fairly large (>0.75). 

Find large correlations for `raw` or `adj` and relatively large `logpval` for `p_signal`. Add these to plots of correlations (black circles) and use for pair
plots below.

```{r}
datsub <- dat %>%
  mutate(index = row_number()) %>%
  filter(adj > 0.8 | raw > 0.8,
         logpval > 2) %>%
  filter(adj - raw == max(adj - raw) |
           adj - raw == min(adj - raw))
```

```{r}
rr <- row(cors)[upper.tri(cors)][datsub$index]
cc <- col(cors)[upper.tri(cors)][datsub$index]
traits <- unique(out$trait)
datsub$traitc <- traits[cc]
datsub$traitr <- traits[rr]
datsub <- datsub %>%
  select(traitc, traitr, everything())
datsub %>%
  rename(cor_raw = "raw", cor_adj = "adj") %>%
  select(traitr, pvalr, traitc, pvalc, cor_raw, cor_adj) %>%
  mutate(across(is.numeric, signif, 3))
```

```{r}
p <- ggplot(dat %>%
  arrange(logpval)) +
  aes(raw, adj, color = logpval) +
  geom_point() +
  geom_hline(yintercept = 0.5, col = "darkgray") +
  geom_smooth(col = "black") + 
  facet_wrap(~ adjgrp, scales = "free_x") +
  scale_colour_gradientn(colours=rainbow(4))
```

```{r}
p +
  geom_point(data = datsub, 
    aes(x = raw, y = adj), 
    shape = 21, fill = NA, size = 5, col = "black", stroke = 2)
```

Same plot but coloring by the least significant `p_signal` among a pair of traits.

```{r}
p <- ggplot(dat %>% arrange(logpmax)) +
  aes(raw, adj, color = logpmax) +
  geom_point() +
  geom_hline(yintercept = 0.5, col = "darkgray") +
  geom_smooth(col = "black") + 
  facet_wrap(~ adjgrp, scales = "free_x") +
  scale_colour_gradientn(colours=rainbow(4))
```

```{r}
p +
  geom_point(data = datsub, 
    aes(x = raw, y = adj), 
    shape = 21, fill = NA, size = 5, col = "black", stroke = 2)
```

```{r}
filter_best <- function(out, traitc, traitr) {
  out %>%
    filter(trait %in% c(traitc, traitr)) %>%
    pivot_longer(signal:mean, names_to = "type", values_to = "value") %>%
    pivot_wider(names_from = "trait", values_from = "value") %>%
    rename(x = traitc,
           y = traitr)
}
```

```{r}
plot_best <- function(out1, summary_type, datsub) {
  ggplot(out1 %>% filter(type == summary_type)) +
    aes(x, y, fill = strain) +
    geom_smooth(method = "lm", se = FALSE, fill = "white", col = "darkgray", formula = "y ~ x") +
    geom_point(
      ggplot2::aes(shape = sex), size = 3, color = "black", alpha = 0.65) +
    ggplot2::scale_shape_manual(values = c(23, 22)) +
    ggplot2::scale_fill_manual(values = CCcolors) +
    facet_wrap(~ sex + diet) +
    xlab(datsub$traitc) +
    ylab(datsub$traitr) +
    ggtitle(paste(
      summary_type, "for",
      "raw cor", round(datsub$raw, 2),
      ", adj cor", round(datsub$adj, 2),
      ", logpvals", 
      paste(round(c(datsub$pvalr, datsub$pvalc), 2), collapse = ", "))) +
    theme(legend.position = "none")
}
```

```{r}
out1 <- filter_best(out, datsub$traitc[1], datsub$traitr[1])
plot_best(out1, "mean", datsub[1,])
```

```{r}
plot_best(out1, "signal", datsub[1,])
```

```{r}
out2 <- filter_best(out, datsub$traitc[2], datsub$traitr[2])
plot_best(out2, "mean", datsub[2,])
```

```{r}
plot_best(out2, "signal", datsub[2,])
```


