---
title: "Founder Diet Study Data"
author: "Brian Yandell"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
params:
  dataset:
    label: "Measurement Set"
    value: PlaMet0
    input: select
    choices: [physio, plasma, liver]
  traits:
    label: "Number of traits"
    value: 3
    input: slider
    min: 2
    max: 6
    
---

```{r setup, include=FALSE, echo = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                      comment = "", fig.height = 7, fig.width = 7)
```

**NB: This document is out of date.**

```{r}
library(tidyverse)
library(readxl)
library(readr)
library(foundr) # github/byandell/foundr
```

```{r}
traitData <- readRDS("plasmaData.rds")
```

# Volcano Plot Ideas

```{r}
dat <- traitData %>% filter(trait == trait[1])
datb <- strainstats(dat)
```

```{r}
signal <- "strain * sex * condition"
form <- stats::formula(paste("value", "~", signal))
fit <- stats::lm(form, dat)
drops <- broom::tidy(drop1(fit,fit, test = "F"))
sdcomp <- drops$sumsq[-1] / drops$df[-1]
names(sdcomp) <- drops$term[-1]
sqrt(sdcomp)
```
Signed coefficients. All others involve multiple strains.

```{r}
coef(fit)[c("sexM", "conditionHF_LC", "sexM:conditionHF_LC")]
```

# Plotly of traits

```{r}
meas <- readRDS("traitData.rds") %>%
  filter(dataset %in% params$dataset)
traitStats <- readRDS("traitStats.rds") %>%
  filter(dataset %in% params$dataset)
```

```{r}
traits <- unique(traitStats$trait)[seq_len(as.integer(params$traits))]
meas <- meas %>% 
  filter(trait %in% traits)
```

Here we pick out 6 compounds with the lowest absolute value and compares sex, condition and strain.

```{r}
plot(traitSolos(meas))
```

```{r warnings = FALSE}
condition <- "sex_condition"
form <- formula(".data$trait ~ .data$strain")
ggplot(meas %>%
         unite(sex_condition, sex, condition)) +
  aes(sex_condition, value, col = .data[[condition]]) +
  geom_jitter() +
  facet_grid(form, scales = "free_y") +
  theme(legend.position = "none", axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
```

```{r}
dat <- meas %>%
  filter(trait %in% traits[1:2]) %>%
  select(strain, animal, sex, condition, trait, value) %>%
  pivot_wider(names_from = "trait", values_from = "value")
plotly::ggplotly(
  ggplot(dat) +
    aes(.data[[traits[1]]], .data[[traits[2]]], color = strain) +
    geom_smooth(method = "lm", se = FALSE) +
    geom_point(size = 2) +
    scale_color_manual(values = CCcolors) +
    facet_grid(sex ~ condition, scales = "free")) 
```

## More than two pairs

This works for 3 but gets crowded for more

```{r}
traitpairs <- as.data.frame(combn(traits, 2))
dat <- 
  bind_rows(
    map(
      traitpairs,
      function(x) {
        meas %>%
          filter(trait %in% x) %>%
          mutate(trait = c("left","right")[match(trait, x)]) %>%
          select(strain, animal, sex, condition, trait, value) %>%
          pivot_wider(names_from = "trait", values_from = "value")
      }),
    .id = "pair") %>%
  unite(sex_condition, sex, condition) %>%
  mutate(pair1 = as.matrix(traitpairs)[1, pair],
         pair2 = as.matrix(traitpairs)[2, pair]) %>%
  select(pair1, pair2, everything())

plotly::ggplotly(
  ggplot(dat) +
    aes(left, right, color = strain) +
    geom_smooth(method = "lm", se = FALSE) +
    geom_point(size = 2) +
    scale_color_manual(values = CCcolors) +
    facet_grid(pair2 + pair1 ~ sex_condition, scales = "free"))
```

```{r}
traitpairs2 <- traitpairs %>%
  mutate(across(
    everything(), 
    function(x) {
      c(paste(x, collapse = ":"),
      paste(rev(x), collapse = ":"))
      })) %>%
  unlist()
```

```{r}
dat <- 
  map(
    traitpairs2,
    function(x) {
      x <- str_split(x, ":")[[1]]
      out <- meas %>%
        filter(trait %in% x) %>%
        mutate(trait = c(x[1],x[2])[match(trait, x)]) %>%
        select(strain, animal, sex, condition, trait, value) %>%
        pivot_wider(names_from = "trait", values_from = "value") %>%
        unite(sex_condition, sex, condition)
      ggplot(out) +
        aes(.data[[x[1]]], .data[[x[2]]], color = strain) +
        geom_smooth(method = "lm", se = FALSE, formula = "y ~ x") +
        geom_point(size = 2) +
        scale_color_manual(values = CCcolors) +
        facet_grid(. ~ sex_condition) +
        theme(legend.position = "none")
    })
```


```{r}
tmp <- names(dat)[1:3]
patchwork::wrap_plots(dat[tmp], nrow = length(tmp))
```

## Other ideas

Replaced variability in order (which does not really work) with
`p_signal` something like `broom::tidy(stats::anova(fitfull, fitred))`
where `fitfull` has model 

`trait ~ strain * sex * condition`

and `fitred` has model

`trait ~ strain * se + sex * condition`

That is, we remove overall strain effect but combine `strain:sex`, `strain:condition` and `strain:sex:condition`.

```{r}
tmp <- meas %>% filter(trait == traits[1])
formful <- formula(value ~ strain * sex * condition)
fitful <- stats::lm(formful, tmp)
formred <- formula(value ~ strain * sex + sex * condition)
fitred <- stats::lm(formred, tmp)
      
(broom::tidy(stats::anova(fitred, fitful)) %>%
  select(p.value))[2,]$p.value
```

```{r}
rr <- resid(fitred)
```

### p-value exploration

```{r}
traitStats <- readRDS("traitStats.rds")
```

```{r}
logpval <- function(x) pmin(5, -log10(x))
ggplot(traitStats %>%
         select(-SD) %>%
         filter(term %in%
                  c("signal", "strain_condition", "strain_sex_condition")) %>%
         mutate(p.value = logpval(p.value)) %>%
         pivot_wider(names_from = "term", values_from = "p.value")) +
  aes(strain_condition, strain_sex_condition, col = signal) +
  geom_point(alpha = 0.75, shape = 21) +
  geom_smooth() +
  facet_wrap(~ dataset, scales = "free") +
  scale_colour_gradientn(colours=rainbow(4))
```

```{r}
ggplot(traitStats %>% filter(term != "noise")) +
  aes(-log10(p.value), col = dataset) +
  geom_density() +
  facet_wrap(~ term) +
  scale_x_log10()
```

# Analysis over time

```{r}
week_traits <- timetraits(traitStats, "week")
min_traits <- timetraits(traitStats, "minute")
```

```{r}
traitSignal <- readRDS("traitSignal.rds")
```

```{r}
minData <- traitTimes(traitData, traitSignal, min_traits[1:2], "cellmean", "minute")
```

```{r}
plot(minData)
```

```{r}
plot(minData, facet_strain = TRUE)
```

## Stats over time

Pieces are in place, a little kludgey but externally looks good. Think about how to install on app.

```{r}
minStats <- traitTimes(traitStats, min_traits[11:13], "p.value", "minute_summary")
```

```{r}
plot(minStats)
```

```{r}
minStats <- traitTimes(traitStats, min_traits[11:13], "p.value", "minute")
```

```{r}
plot(minStats)
```

```{r}
minStats <- traitTimes(traitStats, min_traits[1:5], "p.value", "minute_summary")
```

```{r}
plot(minStats)
```

```{r}
minStats <- traitTimes(traitStats, min_traits[1:5], "p.value", "minute_summary")
```

```{r}
plot(minStats)
```

```{r}
minStats <- traitTimes(traitStats, min_traits[1:5], "p.value", "minute")
```

```{r}
plot(minStats)
```

```{r}
minStats <- traitTimes(traitStats, min_traits[6:10], "p.value", "minute")
```

```{r}
plot(minStats)
```

## Body Weight (BW)

```{r}
wkData <- traitTimes(traitData, traitSignal, week_traits[1], "cellmean", "week")
```

```{r}
plot(wkData)
```

```{r}
wkData <- traitTimes(traitStats, traitnames = week_traits[1], response = "p.value", timecol = "week")
```

```{r}
plot(wkData)
```

```{r}
wkData <- traitTimes(traitStats, traitnames = week_traits[1], response = "SD", timecol = "week")
```

```{r}
plot(wkData)
```

```{r}
bwsum <- PhysioStats %>%
  filter(trait %in% bw, grepl("wk$", trait)) %>%
  mutate(trait = as.numeric(str_remove(str_remove(trait, "^BW_"), "wk$"))) %>%
  rename(week = "trait") %>%
  select(-dataset, -strain_sex) %>%
  pivot_longer(-week, names_to = "term", values_to = "pvalue") %>%
  mutate(pvalue = -log10(pvalue)) %>%
  arrange(term, week)
```

```{r}
ggplot(bwsum) +
  aes(week, pvalue, col = term) +
  geom_point() +
  geom_path()
```

```{r}
outbw <- out %>%
  filter(trait %in% bw, grepl("wk$", trait)) %>%
  mutate(trait = as.numeric(str_remove(str_remove(trait, "^BW_"), "wk$"))) %>%
  rename(week = "trait")
```

```{r}
ggplot_time(outbw)
```

```{r}
ggplot_time(outbw %>% select(-value, -ancillary, -noise) %>% rename(value = "signal"))
```

## Kcal consumed

```{r}
kcsum <- PhysioStats %>%
  filter(trait %in% consumed, grepl("^kcal_", trait)) %>%
  mutate(trait = as.numeric(str_remove(str_remove(trait, "^kcal_consumed_"), "wk$"))) %>%
  rename(week = "trait") %>%
  select(-dataset, -rawSD, -strain.sex) %>%
  pivot_longer(-week, names_to = "term", values_to = "pvalue") %>%
  mutate(pvalue = -log10(pvalue)) %>%
  arrange(term, week)
```

```{r}
ggplot(kcsum) +
  aes(week, pvalue, col = term) +
  geom_point() +
  geom_path()
```


```{r}
outkcal <- out %>%
  filter(trait %in% consumed, grepl("^kcal_", trait)) %>%
  mutate(trait = as.numeric(str_remove(str_remove(trait, "^kcal_consumed_"), "wk$"))) %>%
  rename(week = "trait")
```

```{r}
ggplot_time(outkcal)
```

```{r}
ggplot_time(outkcal %>% select(-value, -ancillary, -noise) %>% rename(value = "signal"))
```

## GTT consumed

```{r}
gttsum <- PhysioStats %>%
  filter(trait %in% gtt, grepl("^GTT_", trait)) %>%
  mutate(trait = str_remove(str_remove(trait, "^GTT_"), "wk$"),
         week = as.numeric(str_remove(trait, ".*_")),
         dose = as.numeric(str_remove(str_remove(trait, "^[a-z]*_"), "_[0-9]*$")),
         trait = str_remove(trait, "_.*$")) %>%
  select(-dataset, -rawSD, -strain.sex) %>%
  select(trait, week, dose, everything()) %>%
  pivot_longer(-(trait:dose), names_to = "term", values_to = "pvalue") %>%
  mutate(pvalue = -log10(pvalue)) %>%
  arrange(term, dose, week)
```

```{r}
ggplot(gttsum %>% filter(trait == "glu")) +
    aes(dose, pvalue, col = term) +
    geom_point(size = 2) +
    geom_smooth(se = FALSE) +
    facet_grid(. ~ week) +
  ggtitle("GTT glucose log10 pvalues")
```

```{r}
ggplot(gttsum %>% filter(trait != "glu")) +
    aes(dose, pvalue, col = term) +
    geom_point(size = 2) +
    geom_smooth(se = FALSE) +
    facet_wrap(~ trait) +
  ggtitle("GTT cpep and ins log10 pvalues")
```


```{r}
outgtt <- out %>%
  filter(trait %in% gtt, grepl("^GTT_", trait)) %>%
  mutate(trait = str_remove(str_remove(trait, "^GTT_"), "wk$"),
         week = as.numeric(str_remove(trait, ".*_")),
         dose = as.numeric(str_remove(str_remove(trait, "^[a-z]*_"), "_[0-9]*$")),
         trait = str_remove(trait, "_.*$"))
```

```{r}
ggplot(outgtt %>% filter(trait == "glu") %>% unite("sex_condition", sex, condition)) +
    aes(dose, value, fill = strain, color = strain) +
    geom_jitter(shape = 21, size = 2, color = "black") +
    geom_smooth(se = FALSE) +
    facet_grid(week ~ sex_condition) +
    scale_color_manual(values = CCcolors) +
    scale_fill_manual(values = CCcolors)
```

```{r}
ggplot(outgtt %>% select(-value, -ancillary, -noise) %>% rename(value = "signal") %>%
         filter(trait == "glu") %>% unite("sex_condition", sex, condition)) +
    aes(dose, value, fill = strain, color = strain) +
    geom_jitter(shape = 21, size = 2, color = "black") +
    geom_smooth(se = FALSE) +
    facet_grid(week ~ sex_condition) +
    scale_color_manual(values = CCcolors) +
    scale_fill_manual(values = CCcolors)
```

```{r}
ggplot(outgtt %>% filter(trait != "glu") %>% unite("sex_condition", sex, condition)) +
    aes(dose, value, fill = strain, color = strain) +
    geom_jitter(shape = 21, size = 2, color = "black") +
    geom_smooth(se = FALSE) +
    facet_grid(trait ~ sex_condition, scales = "free_y") +
    scale_color_manual(values = CCcolors) +
    scale_fill_manual(values = CCcolors)
```

```{r}
ggplot(outgtt %>% select(-value, -ancillary, -noise) %>% rename(value = "signal")
       %>% filter(trait != "glu") %>% unite("sex_condition", sex, condition)) +
    aes(dose, value, fill = strain, color = strain) +
    geom_jitter(shape = 21, size = 2, color = "black") +
    geom_smooth(se = FALSE) +
    facet_grid(trait ~ sex_condition, scales = "free_y") +
    scale_color_manual(values = CCcolors) +
    scale_fill_manual(values = CCcolors)
```


# Clustering and other methods for dimension reduction of multiple traits

WCGNA / clustering to find similar traits.

Strategy:

- use all data or prescreen in some way (based on p_signal?)
- group by names/types (for physio) or automate
- how to find representer for groups of traits
- making sense of representer back to traits?

Operations:

- raw vs adj values
- use `scale()` to standardized a matrix with data in columns
- create some `dist()` object, probably using "euclidean" metric
- decide on number of clusters
- find representer (eigenvector) for clusters

**need to add how many traits in each group**

```{r}
knitr::knit_exit()
```

#### Clustering for physio

```{r}
d <- 1 - cors^2
fit <- hclust(as.dist(d), method="ward.D2")
plot(fit) # display dendogram
```


```{r}
d <- 1 - corsraw^2
fit <- hclust(as.dist(d), method="ward.D2")
plot(fit) # display dendogram
```


#### Biplot

Physio traits

```{r}
outr <- out %>% 
  select(-value, -ancillary, -noise) %>%
  rename(value = "signal")
```

Notice two major axes for consumed.

```{r}
biggplot(biplot_pca(biplot_data(outr, consumed))) +
  ggtitle("consumed")
```

Less clear signal for body weight.

```{r}
biggplot(biplot_pca(biplot_data(outr, bw))) +
  ggtitle("Body Weight")
```


The following shows GTT measures for insulin and glucose are orthogonal.

```{r}
biggplot(biplot_pca(biplot_data(outr, gtt))) +
  ggtitle("GTT")
```

Percent variation by PC component.

```{r}
biplot_pca(biplot_data(outr, gtt)) -> tmp
100* tmp$sdev^2/sum(tmp$sdev^2)
```

factor loadings

```{r}
round(tmp$loadings[,1:3], 2)
```

## Clustering for plasma

```{r}
plasmasum <- readRDS("traitStats.rds") %>%
  filter(dataset == "plasma",
         p_signal < 0.01) # restrict to traits with strong strain*condition signals
```

```{r}
outma <- partition(readRDS("plasma.rds")) %>%
  select(names(out)) %>%
  filter(trait %in% plasmasum$trait)
```

Spearman rank correlation of residuals after removing `strain*sex` and `condition*sex` effects.
That is, looking at correlation due to `strain*condition` and `strain*condition*sex` effects only.

```{r}
corsmaraw <- abscorcalc(outma %>% select(-signal, -ancillary, -noise))
corsma <- abscorcalc(outma %>% select(-value, -ancillary, -noise) %>% rename(value = "signal"))
```

```{r}
d <- 1 - corsma^2
fit <- hclust(as.dist(d), method="ward.D2")
plot(fit) # display dendogram
```

```{r}
groups <- cutree(fit, 4)
```

```{r}
outmar <- outma %>% 
  select(-value, -ancillary, -noise) %>%
  rename(value = "signal")
```

```{r}
pca <- list()
for(i in 1:4) {
  pca[[i]] <- biplot_pca(biplot_data(outmar, names(groups[groups == i])))
}
```

```{r}
tmp <- as.data.frame(lapply(pca, function(x) round(100* (x$sdev^2)[1:3] / sum(x$sdev^2), 2)))
names(tmp) <- seq_along(tmp)
tmp
```

```{r}
p <- list()
for(i in 1:4) {
  p[[i]] <- biggplot(pca[[i]]) +
    ggtitle(paste("plasma group", i))
}
```

```{r fig.length = 7}
patchwork::wrap_plots(p, nrow = length(p))
```

## Clustering for liver

```{r}
liverPvalue <- readRDS("traitStats.rds") %>%
  filter(dataset == "liver",
         p_signal < 0.0001) # restrict to traits with strong strain*condition signals
```

```{r}
liverData <- partition(readRDS("liver.rds")) %>%
  select(names(out)) %>%
  filter(trait %in% liverPvalue$trait)
```

Spearman rank correlation of residuals after removing `strain*sex` and `condition*sex` effects.
That is, looking at correlation due to `strain*condition` and `strain*condition*sex` effects only.

```{r}
corsli <- abscorcalc(liverData %>% select(-value, -ancillary, -noise) %>% rename(value = "signal"))
```

```{r}
d <- 1 - corsli^2
fit <- hclust(as.dist(d), method="ward.D2")
plot(fit) # display dendogram
```

```{r}
groups <- cutree(fit, 4)
```

```{r}
liverDatar <- liverData %>% 
  select(-value, -ancillary, -noise) %>%
  rename(value = "signal")
```

```{r}
pcali <- list()
for(i in 1:4) {
  pcali[[i]] <- biplot_pca(biplot_data(liverDatar, names(groups[groups == i])))
}
```

```{r}
tmp <- as.data.frame(lapply(pcali, function(x) round(100* (x$sdev^2)[1:3] / sum(x$sdev^2), 2)))
names(tmp) <- seq_along(tmp)
tmp
```

```{r}
p <- list()
for(i in seq_along(pcali)) {
  p[[i]] <- biggplot(pcali[[i]]) +
    ggtitle(paste("plasma group", i))
}
```

```{r fig.length = 7}
patchwork::wrap_plots(p, nrow = length(p))
```

## Manova?

