---
title: "Founder Diet Study Data"
author: "Brian Yandell"
date: "2022-12-05"
output: html_document
---

```{r setup, include=FALSE, echo = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                      comment = "", fig.width = 7, fig.height = 7)
```
 
See file `X202SC22070248-Z01-F001_Report.html` for QC Analysis Report on liver RNAseq data.

```{r}
library(tidyverse)
library(readxl)
library(readr)
```

# Mouse annotations

```{r}
excel_sheets("data/mouse annotations for founder diet study.xlsx")
annot <- read_excel("data/mouse annotations for founder diet study.xlsx") %>%
  mutate(diet = ifelse(as.character(diet_no) == "200339", "HC_LF", "HF_LC"))
```

```{r}
annot %>%
  distinct(diet, diet_no, diet_description)
```

```{r}
annot %>%
  count(strain, sex, diet) %>%
  unite("sex_diet", sex, diet) %>%
  pivot_wider(names_from = "sex_diet", values_from = "n")
```

# Physiological trait counts by strain, sex, diet

```{r}
physio <- readRDS("physio.rds")
```

There are `r nrow(distinct(physio, trait))` distinct traits.

Count of mice per strain, sex and diet. All combinations have at least some measurements.

```{r}
physio %>%
  filter(!is.na(value)) %>%
  distinct(strain, number, sex, diet) %>%
  count(strain, sex, diet) %>%
  unite("sex_diet", sex, diet) %>%
  pivot_wider(names_from = "sex_diet", values_from = "n")
```

Percent missing values across traits

```{r}
physio %>%
  filter(!is.na(value)) %>%
  count(strain, sex, diet) %>%
  unite("sex_diet", sex, diet) %>%
  mutate(n = round(100 * (1 - n / max(n)), 2)) %>%
  pivot_wider(names_from = "sex_diet", values_from = "n")
```


CAST is only measured on `r nrow(distinct(filter(physio, !is.na(value), strain == "CAST"), trait))` traits,
but those all have 6 measurements per diet and sex.

```{r}
cast_traits <- physio %>%
  filter(!is.na(value), strain == "CAST") %>%
  count(trait)
range(cast_traits$n)
cast_traits <- cast_traits$trait
```

NZO males on diet 200339 had half missing values. Of those with non-missing values, most had all 6 and a few had only 5.

```{r}
nzo_traits <- physio %>%
  filter(!is.na(value), strain == "NZO", sex == "M", diet == "200339") %>%
  count(trait) %>%
  filter(n > 0)
nzo_traits %>%
  count(n, name = "traits")
nzo_traits <- nzo_traits$trait
```

All the NZO traits with males on diet 200339 (`r sum(!is.na(match(nzo_traits, cast_traits)))`) have data for CAST mice.

# Plasma metabolite abundance data

```{r}
plasma <- readRDS("plasma.rds")
```

There are missing data for B6 mice across 3 of the 4 sex:diet combinations.

```{r}
plasma %>%
  filter(!is.na(value)) %>%
  distinct(strain, number, sex, diet) %>%
  count(strain, sex, diet) %>%
  unite("sex_diet", sex, diet) %>%
  pivot_wider(names_from = "sex_diet", values_from = "n")
```
All of the `r nrow(distinct(filter(plasma, !is.na(value), strain == "B6"), number))` B6 mice with non-missing data have data on all the compounds.

```{r}
b6_plasma <- plasma %>%
  filter(!is.na(value), strain == "B6") %>%
  count(trait)
range(b6_plasma$n)
b6_plasma <- b6_plasma$trait
```

Now some plots of plasma data. It is difficult to tell much

```{r warnings = FALSE}
ggplot(plasma) +
  aes(value, col = strain) +
  geom_density() +
  facet_grid(diet ~ sex) +
  theme(legend.position = "none")
```

This picks out 6 compounds with the highest absolute value and compares sex, diet and strain.

```{r}
plasmamax <- plasma %>%
  group_by(trait) %>%
  summarize(maxplas = max(abs(value)),
            minplas = min(abs(value)),
            varplas = var(value),
            .groups = "drop") %>%
  ungroup() %>%
  arrange(desc(maxplas))
```

```{r warnings = FALSE}
tmp <- plasmamax$trait[seq(6)]
ggplot(plasma %>%
         filter(trait %in% tmp) %>%
         unite(sex_diet, sex, diet)) +
  aes(sex_diet, value, col = sex_diet) +
  geom_jitter() +
  facet_grid(trait ~ strain, scales = "free_y") +
  theme(legend.position = "none", axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
```

Here we pick out 6 compounds with the lowest absolute value and compares sex, diet and strain.

```{r warnings = FALSE}
tmp <- (plasmamax %>% arrange(minplas))$trait[seq(6)]
ggplot(plasma %>%
         filter(trait %in% tmp) %>%
         unite(sex_diet, sex, diet)) +
  aes(sex_diet, value, col = sex_diet) +
  geom_jitter() +
  facet_grid(trait ~ strain, scales = "free_y") +
  theme(legend.position = "none", axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
```

Here are the 6 with most variation.

```{r warnings = FALSE}
tmp <- (plasmamax %>% arrange(desc(varplas)))$trait[seq(6)]
ggplot(plasma %>%
         filter(trait %in% tmp) %>%
         unite(sex_diet, sex, diet)) +
  aes(sex_diet, value, col = sex_diet) +
  geom_jitter() +
  facet_grid(trait ~ strain, scales = "free_y") +
  theme(legend.position = "none", axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
```

### Exploratory Scatterplots for Plasma

```{r}
library(GGally)
tmp <- (plasmamax %>% arrange(desc(varplas)))$trait[seq(6)]
tmp2 <- plasma %>%
        filter(trait %in% tmp[c(3,6)]) %>%
        select(strain, number, sex, diet, trait, value) %>%
        pivot_wider(names_from = "trait", values_from = "value")
plotly::ggplotly(ggplot(tmp2) +
  aes(.data[[tmp[3]]], .data[[tmp[6]]], color = strain) +
  geom_point(size = 2) +
  scale_color_manual(values = CCcolors) +
  facet_grid(sex ~ diet))
```

# Liver RNAseq

Note that annotation file relates the number to the sex and diet.

```{r}
liver <- readRDS("liver.rds")
```

Apparently no missing values.

```{r}
liver %>%
  filter(!is.na(value)) %>%
  count(strain, number) %>%
  filter(n < n[1])
```

But there are a lot of small values. Here are densities by strain, with separate lines
by mouse number.

```{r}
ggplot(liver) +
  aes(value, col = number) +
  geom_density() +
  scale_x_log10() +
  facet_wrap(~ strain) +
  theme(legend.position = "none")
```

A quick examination of a few rna sequences. Each colored line has the 24 measurements
for that strain.

```{r}
livermax <- liver %>%
  group_by(trait) %>%
  summarize(maxtrait = max(value),
            vartrait = var(value),
            .groups = "drop") %>%
  ungroup() %>%
  arrange(desc(maxtrait))
```

Here are six with the highest values.

```{r}
tmp <- livermax$trait[seq(6)]
ggplot(liver %>%
         filter(trait %in% tmp) %>%
         unite(sex_diet, sex, diet)) +
  aes(sex_diet, value, col = sex_diet) +
  geom_jitter() +
  facet_grid(trait ~ strain, scales = "free_y") +
  theme(legend.position = "none", axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
```

Here are 6 with the highest variance

```{r}
tmp <- (livermax %>% arrange(desc(vartrait)))$trait[seq(6)]
ggplot(liver %>%
         filter(trait %in% tmp) %>%
         unite(sex_diet, sex, diet)) +
  aes(sex_diet, value, col = sex_diet) +
  geom_jitter() +
  facet_grid(trait ~ strain, scales = "free_y") +
  theme(legend.position = "none", axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
```

Here are the 6 with the highest variance.

```{r}
tmp <- unique((liver %>% filter(value > 19))$trait)[seq(6)]
ggplot(liver %>%
         filter(trait %in% tmp) %>%
         unite(sex_diet, sex, diet)) +
  aes(sex_diet, value, col = sex_diet) +
  geom_jitter() +
  facet_grid(trait ~ strain, scales = "free_y") +
  theme(legend.position = "none", axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
```
