---
title: "WGCNA"
author: "Brian Yandell"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
options(stringsAsFactors = FALSE)
```

```{r eval = FALSE}
# Need VPN and connection to ResearchDrive server

RD <- "/Volumes/adattie/General"
RDdir <- file.path(RD, "PersonalSharedFolder2/Mark/DO_multitissue")
list.files(RDdir, "*.R$")
```

```{r eval = FALSE}
if (!require("BiocManager", quietly = TRUE))
  install.packages("BiocManager")
BiocManager::install("GO.db")
BiocManager::install("impute")
BiocManager::install("preprocessCore")
```

```{r}
datadir <- "~/Documents/Research/attie_alan/FounderDietStudy"
setwd(datadir)
datadir <- "."
```

# WGCNA 

Read or save (RDS) module information for each dataset.

```{r}
datasets <- c("LivMet","PlaMet0","PlaMet120","LivRna")
```

```{r}
for(mod in datasets) {
  modname <- paste0(mod, "Module")
  if(file.exists(
    filename <- file.path(datadir, paste0(mod, "Module.rds")))) {
    assign(modname, readRDS(filename))
  } else {
    # Create list of WGCNA modules.
    assign(modname, 
      modulr::listof_wgcnaModules(
        readRDS(file.path(datadir, paste0(mod, "Data", ".rds"))),
        readRDS(file.path(datadir, paste0(mod, "Signal", ".rds")))))
    
    # Save listof_wgcnaModules object.
    saveRDS(
      get(modname),
      filename)
  }
}
```

Create `traitModule` object as list of dataset module objects.

```{r}
if(file.exists(filename <- file.path(datadir, "traitModule.rds"))) {
  traitModule <- readRDS(filename)
} else {
  traitModule <- list()
  for(mod in datasets) {
    traitModule[[mod]] <- 
      readRDS(
        file.path(
          datadir,
          paste0(mod, "Module.rds")))
  }
  saveRDS(traitModule, filename)
}
```

```{r}
summary(traitModule[["LivMet"]], "signal")
```

# Module Harmony

```{r}
moduleData <- 
  dplyr::mutate(
    foundr::moduleHarmony("Module",, traitModule, response = "value"),
    dataset = "Module")
```

```{r}
moduleStats <- foundr::strainstats(moduleData)
```

```{r}
moduleSignal <- foundr::partition(moduleData)
```

```{r}
(modstats <- dplyr::arrange(
  tidyr::pivot_wider(
  dplyr::mutate(
    dplyr::select(
      dplyr::filter(
        moduleStats,
        !(trait %in% c("grey")),
        term %in% c("signal","cellmean","rest")),
      -SD),
    p.value = signif(-log10(p.value), 3)),
  names_from = "term", values_from = "p.value"),
dplyr::desc(signal)))
```

```{r}
(traits <-
   dplyr::filter(
     tidyr::unite(
       modstats,
       datatraits,
       dataset, trait,
       sep = ": "),
     signal > 1)$datatraits)
```

```{r}
moduleSolo <- foundr::traitSolos(
  moduleData, moduleSignal, response = "cellmean",
  traitnames = traits[1:4])
```

Cell mean for top LivMet and PlaMet module

```{r}
foundr::ggplot_traitSolos(moduleSolo)
```

```{r}
moduleSolo <- foundr::traitSolos(
  moduleData, moduleSignal, response = "signal",
  traitnames = traits[1:4])
```

Signal for top LivMet and PlaMet module

```{r}
foundr::ggplot_traitSolos(moduleSolo)
```

```{r}
traitData <- readRDS(file.path(datadir, "traitData.rds"))
traitSignal <- readRDS(file.path(datadir, "traitSignal.rds"))
```

### LivMet greenyellow module

```{r}
(GYLivMet <- 
  dplyr::arrange(
    dplyr::filter(
      traitModule$LivMet$value$modules,
      module == "greenyellow"),
    dplyr::desc(abs(kME))))
```

```{r}
GYSolo <- foundr::traitSolos(
  traitData, traitSignal, response = "cellmean",
  traitnames = paste("LivMet", GYLivMet$trait[1], sep = ": "))
```

```{r}
foundr::ggplot_traitSolos(GYSolo)
```

```{r}
foundr::ggplot_traitSolos(GYSolo, facet_strain = TRUE)
```

### LivMet black module

```{r}
(BlLivMet <- 
  dplyr::arrange(
    dplyr::filter(
      traitModule$LivMet$value$modules,
      module == "black"),
    dplyr::desc(abs(kME))))
```

```{r}
BlSolo <- foundr::traitSolos(
  traitData, traitSignal, response = "cellmean",
  traitnames = paste("LivMet", BlLivMet$trait[2], sep = ": "))
```

```{r}
foundr::ggplot_traitSolos(GYSolo)
```

```{r}
BlSolo <- foundr::traitSolos(
  traitData, traitSignal, response = "signal",
  traitnames = paste("LivMet", BlLivMet$trait[2], sep = ": "))
```

```{r}
foundr::ggplot_traitSolos(GYSolo)
```

## Plots

```{r}
modband <- foundr::module_band(traitModule$LivMet, "signal")
```

```{r}
foundr::ggplot_module_band(modband)
```

```{r warning = FALSE, message = FALSE}
foundr::ggplot_listof_wgcnaModules(traitModule$LivMet)
```
Standard plot.

```{r eval=FALSE}
plot(traitModule$LivMet$value)
```

# Module Comparisons

```{r}
mods <- foundr::module_kMEs(traitModule$LivMet)
```

```{r}
subset(mods, "signal", "cellmean", "blue")
```

```{r}
dplyr::arrange(
    dplyr::count(
      mods,
      value_col),
    dplyr::desc(n))
```


```{r}
ggplot2::autoplot(mods, "value", "cellmean")
```

```{r}
plotly::ggplotly(ggplot2::autoplot(mods, "value", "cellmean", abskME = TRUE,
                  facetmodules = "blue"))
```

# Eigentraits

```{r}
eigens <- foundr::eigen_cor(traitModule$LivMet)
```

```{r}
subset(eigens, "value", "signal")
```

```{r}
ggplot2::autoplot(eigens, "value", "cellmean")
```

The sign of correlation of modules could provide adjustment to direction in kME plots above. However, using the absolute value seems to show the strong relationship (or not).

# Compare Datasets

```{r}
dmods <- list(LivMet = traitModule$LivMet,
              PlaMet0 = traitModule$PlaMet0)
response <- "value"
```

```{r}
valueMet <- foundr::join_dataset_modules(dmods, response)
```

```{r}
metmods <- foundr::module_kMEs(valueMet)
```

```{r}
plotly::ggplotly(
  ggplot2::autoplot(metmods, "LivMet", "PlaMet0"))
```

```{r}
traitData <- 
  dplyr::filter(
      readRDS(file.path(datadir, "traitData.rds")),
      dataset %in% c("LivMet","PlaMet0"))
traitSignal <-
    dplyr::filter(
      readRDS(file.path(datadir, "traitSignal.rds")),
      dataset %in% c("LivMet","PlaMet0"))
```

```{r}
soloData <- 
  foundr::traitSolos(
    dplyr::filter(
      traitData,
      trait == trait[1]),
    dplyr::filter(
      traitSignal,
      trait == trait[1]))
```

```{r}
DT::datatable(soloData)
```

```{r}
foundr::ggplot_traitPairs(
  foundr::traitPairs(soloData))
```

```{r}
eigenMet <- foundr::eigen_cor(valueMet)
```

```{r}
ggplot2::autoplot(eigenMet, "LivMet", "PlaMet0")
```

