---
title: "Module Visualization"
author: "Brian Yandell"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
runtime: shiny
params:
  echo: no
resource_files:
- traitModule.rds
- traitData.rds
- traitSignal.rds
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = params$echo, warning = FALSE, message = FALSE)
```

```{r}
devtools::install_cran("plotly") #  not yet on UW dataviz
devtools::install_cran("markdown") #  not yet on UW dataviz
devtools::install_cran("cowplot") #  not yet on UW dataviz
devtools::install_cran("patchwork") #  not yet on UW dataviz
devtools::install_cran("ggdendro") # not yet in UW dataviz
devtools::install_github("byandell/foundr")
```

```{r}
options(stringsAsFactors = FALSE)
```

This document selects a `dataset` and creates WGCNA modules. Selecting a `response` enables visualization of stats for the modules as well as the WGCNA dendogram and color panel. Selects a `facet response` and `color response` compares two types of WGCNA modules. The first figure plots kMEs while the second one shows correlation of eigentraits across those responses. Data can be downloaded by clicking the button.

### Explanation of Responses

The response for each trait can be partitioned into uncorrelated pieces that
have useful interpretations. Letting `value` be the trait value for an organism, it is partitioned into the `cellmean` and `noise`. The `cellmean` is the part of the response explained by the model, and the `noise` is the residual. The `cellmean` can be decomposed into main effects and interactions involving the factors `strain`, `sex` and `condition`. However, generally we are interested in whether there are different `strain` responses to `condition`, in which case we are interested in the `signal` composed of `strain:condition` and `strain:sex:condition` interactions, with the `rest` of the model components conveniently combined.

<center>
```
value = cellmean + noise
cellmean = signal + rest
value = signal + rest + noise

signal = strain:condition + strain:sex:condition
rest = strain + sex + condition + strain:sex + sex:condition
```
</center>

For each trait, the `cellmean` and `noise` terms are uncorrelated, as are the `signal` and `rest` terms, which add up to the `cellmean`. That means their variances add up ($V_i = V_s + V_r + V_n$ and $V_c = V_s + V_r$), and the covariances of uncorrelated terms are 0 ($C_{cn} = 0, C_{sr} = 0$). This makes it possible later to compare across different datasets in interesting ways.

Patterns emerge when comparing these components across traits within WGCNA modules. That is, a module identified with the `indivdual` response (the default approach) may represent primarily `signal`, which would be useful. However, if a module is primarily associated with `rest` or `noise`, it is less likely to yield interesting trait relationships. Further, comparing modules across responses might identify interesting subsets of modules for further study. 

```{r}
datadir <- "."
```

```{r}
datasets <- c("LivMet","PlaMet")
responses <- c("value","cellmean","signal","rest","noise")
```

```{r}
traitModule <- readRDS(file.path(datadir, "traitModule.rds"))
```

```{r}
traitData <- 
  dplyr::filter(
      readRDS(file.path(datadir, "traitData.rds")),
      dataset %in% datasets)
traitSignal <-
    dplyr::filter(
      readRDS(file.path(datadir, "traitSignal.rds")),
      dataset %in% datasets)
```

```{r}
shiny::fluidRow(
  shiny::column(
    6,
    shiny::selectInput(
      "dataset", "Dataset:",
      datasets)),
  shiny::column(
    6,
    shiny::selectInput(
      "response", "Response:",
      responses))
)
```

# WGCNA 

```{r}
shiny::renderPlot({
  shiny::req(input$dataset, input$response)
  foundr::ggplot_listof_wgcnaModules(traitModule[[input$dataset]], input$response)
})
```

```{r}
DT::renderDataTable({
  shiny::req(input$dataset, input$response)
  DT::datatable(
    dplyr::filter(
      summary(traitModule[[input$dataset]]),
      response == input$response),
    options = list(paging =TRUE, pageLength = 5))
})
```

# Module Comparisons

```{r}
mods <- shiny::reactive({
  shiny::req(input$dataset)
  foundr::module_kMEs(traitModule[[input$dataset]])
})
```

```{r}
fmodules <- shiny::reactive({
  shiny::req(mods(), input$responseF)
  levels(mods()[[paste0(input$responseF, "_col")]])
})
```

```{r}
cmodules <- shiny::reactive({
  shiny::req(mods(), input$responseC)
  levels(mods()[[paste0(input$responseC, "_col")]])
})
```

```{r}
shiny::fluidRow(
  shiny::column(
    4,
    shiny::selectInput(
      "responseF", "Facet Response:",
      "cellmean")),
  shiny::column(
    4,
    shiny::selectInput(
      "responseC", "Color Response:",
      "value"))
)
```

```{r}
shiny::fluidRow(
  shiny::column(
    4,
    shiny::selectInput(
      "fmodules", "Facet Modules:",
      c("gray","turquoise"),
      multiple = TRUE)),
  shiny::column(
    4,
    shiny::selectInput(
      "cmodules", "Color Modules:",
      c("gray","turquoise"),
      multiple = TRUE)),
  shiny::column(
    2,
    checkboxInput("abs", "Absolute kME?"))
)
```

```{r}
responseFs <- shiny::reactive({
  shiny::req(input$response)
  unique(c(input$response, responses))
})
```

```{r}
responseCs <- shiny::reactive({
  shiny::req(input$responseF)
  responses[responses != input$responseF]
})
```

```{r}
shiny::observeEvent(
  input$response,
  {
    shiny::updateSelectInput(
      session, "responseF",
      choices = responseFs(),
      selected = responseFs()[1])
})
```

```{r}
shiny::observeEvent(
  input$responseF,
  {
    shiny::updateSelectInput(
      session, "responseC",
      choices = responseCs(),
      selected = responseCs()[1])

    shiny::updateSelectInput(
      session, "fmodules",
      choices = fmodules(),
      selected = fmodules())
})
```

```{r}
shiny::observeEvent(
  input$responseC,
  {
    shiny::updateSelectInput(
      session, "cmodules",
      choices = cmodules(),
      selected = cmodules())
})
```

```{r}
moddata <- shiny::reactive({
  shiny::req(mods(), input$responseF, input$responseC,
             input$fmodules, input$cmodules)
  
  foundr::subset_module_kMEs(mods(), input$responseF, input$responseC,
         facetmodules = input$fmodules, colormodules = input$cmodules)
})
```

```{r}
plotly::renderPlotly({
  shiny::req(moddata(), input$responseF, input$responseC)
  print(
    ggplot2::autoplot(
      moddata(),
      input$responseF, input$responseC, input$abs))
})
```

```{r}
DT::renderDataTable({
  shiny::req(input$dataset, input$response)
  DT::datatable(
    dplyr::arrange(
      tidyr::pivot_wider(
        dplyr::select(
          summary(traitModule[[input$dataset]]),
          -maxkME, -minkME),
        names_from = "response", values_from = "count",
        values_fill = 0),
      dplyr::desc(.data[[input$response]])),
    options = list(paging =TRUE, pageLength = 5))
})
```

# Eigentrait Correlations

```{r}
eigens <- shiny::reactive({
  shiny::req(input$dataset)
  foundr::eigen_cor(traitModule[[input$dataset]])
})
```

```{r}
plotly::renderPlotly({
  shiny::req(eigens(), input$responseF, input$responseC)
  print(
    ggplot2::autoplot(
      eigens(),
      input$responseF, input$responseC,
      facetmodules = input$fmodules,
      colormodules = input$cmodules))
})
```

The sign of correlation of modules could provide adjustment to direction in kME plots above. However, using the absolute value seems to show the strong relationship (or not).

```{r}
DT::renderDataTable({
  shiny::req(eigens(), input$responseF, input$responseC)
  DT::datatable(
    foundr::subset_eigen_cor(eigens(), input$responseC, input$responseF),
    options = list(paging =TRUE, pageLength = 5))
})
```

# Download Data

```{r}
shiny::downloadHandler(
  filename = "blah.csv",
  content = function(file) {
    req(moddata())
    write.csv(moddata(), file)
  }
)
```

```{r}
DT::renderDataTable({
  shiny::req(moddata())
  dplyr::mutate(
    moddata(),
    dplyr::across(
      dplyr::where(is.numeric),
      function(x) signif(x, 4)))
})
```

# Compare Datasets

```{r}
shiny::fluidRow(
  shiny::column(
    4,
    shiny::selectInput(
      "datasetD", "Facet Dataset:",
      datasets)),
  shiny::column(
    4,
    shiny::selectInput(
      "responseD", "Dataset Response:",
      responses))
)
```

```{r}
shiny::fluidRow(
  shiny::column(
    4,
    shiny::selectInput(
      "fmodules2", "Facet Modules:",
      c("gray","turquoise"),
      multiple = TRUE)),
  shiny::column(
    4,
    shiny::selectInput(
      "cmodules2", "Color Modules:",
      c("gray","turquoise"),
      multiple = TRUE)),
  shiny::column(
    2,
    checkboxInput("absD", "Absolute kME?"))
)
```

```{r}
dataset2 <- shiny::reactive({
  shiny::req(input$datasetD)
  datasets[input$datasetD != datasets][1]
})
```

```{r}
fmodules2 <- shiny::reactive({
  shiny::req(metmods(), input$datasetD)
  levels(metmods()[[paste0(input$datasetD, "_col")]])
})
```

```{r}
cmodules2 <- shiny::reactive({
  shiny::req(metmods(), dataset2())
  levels(metmods()[[paste0(dataset2(), "_col")]])
})
```

```{r}
shiny::observeEvent(
  input$datasetD,
  {
    shiny::updateSelectInput(
      session, "fmodules2",
      choices = fmodules2(),
      selected = fmodules2())
})
```

```{r}
shiny::observeEvent(
  dataset2(),
  {
    shiny::updateSelectInput(
      session, "cmodules2",
      choices = cmodules2(),
      selected = cmodules2())
})
```

```{r}
dmods <- shiny::reactive({
  shiny::req(input$datasetD, dataset2(), input$responseD)

  foundr::join_dataset_modules(
    traitModule[c(input$datasetD, dataset2())],
    input$responseD)
})
```

```{r}
metmods <- shiny::reactive({
  shiny::req(dmods())
  foundr::module_kMEs(dmods())
})
```

```{r}
moddata2 <- shiny::reactive({
  shiny::req(metmods(), input$datasetD, dataset2(),
             input$fmodules2, input$cmodules2)
  
  foundr::subset_module_kMEs(
    metmods(), input$datasetD, dataset2(),
    facetmodules = input$fmodules2,
    colormodules = input$cmodules2)
})
```

```{r}
plotly::renderPlotly({
  shiny::req(moddata2(), input$datasetD, dataset2())
  if(is.null(moddata2()))
    foundr::plot_null("no moddata2 data")
  else {
  plotly::event_register(
    plotly::ggplotly(
      print(
        ggplot2::autoplot(
          moddata2(),
          input$datasetD, dataset2(), input$absD))),
      "plotly_click")
  }
})
```

```{r}
pointKey <- reactive({
  shiny::req(moddata2())
  point <- plotly::event_data(
    event = "plotly_click", priority = "event")
  shiny::req(point) # to avoid error if no point is clicked
  point$key
})
```

```{r}
shiny::renderTable({
  shiny::req(moddata2(), pointKey())
  dplyr::filter(
    moddata2(),
    trait == pointKey()) # use the key to find selected point
})
```

```{r}
pairData <- shiny::reactive({
  shiny::req(pointKey())
  foundr::traitSolos(
    dplyr::filter(
      traitData,
      trait == pointKey()),
    dplyr::filter(
      traitSignal,
      trait == pointKey()))
})
```

```{r}
shiny::renderPlot({
  shiny::req(pairData())
  foundr::ggplot_traitPairs(
    foundr::traitPairs(pairData()))
})
```

```{r}
DT::renderDataTable({
  shiny::req(pairData(), pointKey())
  DT::datatable(
    pairData(),
    options = list(paging =TRUE, pageLength = 5))
})
```

```{r}
eigenMet <- shiny::reactive({
  shiny::req(dmods())
  foundr::eigen_cor(dmods())
})
```

```{r}
plotly::renderPlotly({
  shiny::req(eigenMet(), dataset2(), input$datasetD)
  print(
    ggplot2::autoplot(
      eigenMet(),
      input$datasetD, dataset2(),
      facetmodules = input$fmodules2,
      colormodules = input$cmodules2))
})
```

